Molecular dynamics simulation of synchronization in driven particles

### 2021-01-13 manuscript/manuscript_ajp_2020.tex
removed ramp in F^DC and removed related text

 Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{d}(t) = [F^{dc} + F^{ac} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{dc}$,
and a time dependent component with amplitude $F^{ac}$
and frequency $\omega = 2 \pi f$.
%in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
When $F^{dc}$ is non-zero,  
%In several simulations
we slowly increase 
$F^{dc}$ from zero
to avoid transient behavior
by ensuring the particle is subject to a smoothly varying force.
We illustrate the increase
of $F^{dc}$ in Fig.~\ref{fig:0}(a).

%because
The inertia of 
small particles is greatly reduced by interactions
with fluid particles \cite{Purcell1977}.
Our model 
colloids are overdamped
-- i.e. suspended in a continuous viscous fluid
that dissipates energy at a rate such 
that the particles do not accelerate.
%that greatly impedes their motions.
Energy dissipation from the fluid is modeled
with a frictional force on particle with velocity $\vec{v}_i$
%that removes energy from
%the particles,
%which we model as
$\vec{F}^{drag}_i = -\eta \vec{v}_i$
%proportional to the particle velocity 
with a friction coefficient $\eta$
proportional to the fluid viscosity.
We discuss friction models for
spheres moving through fluids in 
Section~\ref{sec:problems} Exercise~\ref{ex:reynolds}. %for more information
%We model the particle dynamics  with an overdamped equation of motion. Models with inertia set to zero are appropriate for moving through a high viscosity fluid
Newton's second law for an individual particle
%$m \vec{a}_i = \sum \vec{F}$
is simplified
by the assumption $\vec{a}_i$ is zero. % in the overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The overdamped equation of motion for an isolated particle is
%solved for velocity 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i} + \vec{F}^{d}(t).
    \label{eq:motion}
\end{equation}
with  $\eta = 1$.  
%which we describe in more detail in Table~\ref{tab:1}
%and Exercise~\ref{ex:reynolds}.
In %Section~\ref{sec:problems}
Exercise~\ref{ex:n2l}
we invite the reader to confirm this result.

@@

In Fig.~\ref{fig:0} 
the driven particle adjusts
its motion to traverse
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^d(t)$.
%
In Fig.~\ref{fig:0}(a)
we plot $F^d(t)$ %defined in Eq.~\ref{eq:drive}.
as a function of time with
%The time independent
constants 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$ time units.
To avoid transient oscillations at early times, 
the dc driving force is initially zero   
and 
slowly increased %$F^{dc}$.
at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  
to the constant value $F^{dc}=0.1$.
In Fig.~\ref{fig:0}(a) we plot $F^{dc}$
in blue to show the increase in 
$F^{dc}$ for the first $40$ time units
In orange we show 
the combined ac and dc drive $F^d(t)$ 
when the magnitude of the ac applied force is initially zero.
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
(i.e. $\sin{(\omega t)} = 1$).

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^d(t)$ (orange) in Eq.~\ref{eq:drive}
  where the parameter $F^{dc}$ (blue) is slowly increased from zero
  to a constant value
  as described in the text.  (b) 
  The $y-$position of the driven particle
  as a function of time %of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}

>>>>

 Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{d}(t) = [F^{dc} + F^{ac} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{dc}$,
and a time dependent component with amplitude $F^{ac}$
and frequency $\omega = 2 \pi f$.
%in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP!!!
%%%When $F^{dc}$ is non-zero,  we slowly increase $F^{dc}$ from zero to avoid transient behavior by ensuring the particle is subject to a smoothly varying force. We illustrate the increase of $F^{dc}$ in Fig.~\ref{fig:0}(a).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%because
The inertia of 
small particles is reduced by interactions
with fluid particles \cite{Purcell1977}.
Our model 
colloids are overdamped
-- i.e. suspended in a continuous viscous fluid
that dissipates energy at a rate such 
that the particles do not accelerate.
%that greatly impedes their motions.
Energy dissipation from the fluid is modeled
with a frictional force on particle with velocity $\vec{v}_i$
%that removes energy from
%the particles,
%which we model as
$\vec{F}^{drag}_i = -\eta \vec{v}_i$
%proportional to the particle velocity 
with a friction coefficient $\eta$
proportional to the fluid viscosity.
We discuss friction models for
spheres moving through fluids in 
Section~\ref{sec:problems} Exercise~\ref{ex:reynolds}. %for more information
%We model the particle dynamics  with an overdamped equation of motion. Models with inertia set to zero are appropriate for moving through a high viscosity fluid
Newton's second law for an individual particle
%$m \vec{a}_i = \sum \vec{F}$
is simplified
by the assumption $\vec{a}_i$ is zero. % in the overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The overdamped equation of motion for an isolated particle is
%solved for velocity 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i} + \vec{F}^{d}(t).
    \label{eq:motion}
\end{equation}
with  $\eta = 1$.  
%which we describe in more detail in Table~\ref{tab:1}
%and Exercise~\ref{ex:reynolds}.
In %Section~\ref{sec:problems}
Exercise~\ref{ex:n2l}
we invite the reader to confirm this result.

@@

In Fig.~\ref{fig:0} 
the driven particle adjusts
its motion to traverse
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^d(t)$.
%
In Fig.~\ref{fig:0}(a)
we plot $F^d(t)$ %defined in Eq.~\ref{eq:drive}.
as a function of time with
%The time independent
constants $F^{dc}=0.1$, 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$ time units.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP
%To avoid transient oscillations at early times, the dc driving force is initially zero and slowly increased %$F^{dc}$. at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  to the constant value $F^{dc}=0.1$. In Fig.~\ref{fig:0}(a) we plot $F^{dc}$ in blue to show the increase in  $F^{dc}$ for the first $40$ time units In orange we show  the combined ac and dc drive $F^d(t)$ when the magnitude of the ac applied force is initially zero.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
(i.e. $\sin{(\omega t)} = 1$).

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single_particle.pdf}
\caption{(a) The applied driving force $F^d(t)$. % (Eq.~\ref{eq:drive}).
  %where the parameter $F^{dc}$ (blue) is slowly increased from zero %to a constant value %as described in the text.
  (b) 
  The $y-$position of the driven particle
  as a function of time %of a single driven particle
  normalized by the period of the substrate $\lambda$.
  }
\label{fig:0}
\end{figure}
\end{center}
